Ultrafast Electron Microscopes: Design Criteria, Electron Sources, and Column Modeling

### 2013-04-09 inc/photocathode/plasmon/plasmon.tex
Merge remaining comments from andreas on PAPE

Above an incident laser pulse intensity of $\sim$32MW/cm$^2$, there is a rapid and dramatic increase in the emission efficiency for the nano-patterned photocathode.
This can be explained by a suppression of the barrier to photoemission below the two-photon excitation energy, thus allowing direct two-photon photoemission.
This is likely caused by the laser-excited surface plasmon since a similar increase from the neighboring flat gold surface is not present.
The observed distinct changes in the momentum distributions (\ref{fig:pape} B and C) in this laser intensity range support this argument.
In \ref{fig:pape} B, at a laser intensity of $\sim$30MW/cm$^2$, there is already a distinct broadening of the momentum distribution in the direction parallel to the wavevector $k_G$ of the sinusoidal modulation on the photocathode (i.e., perpendicular to the FIB-milled grooves). 
This broadening becomes the dominant feature of the electron emission above $\sim$32MW/cm$^2$; indeed, it becomes so severe that at laser intensities above $\sim$37MW/cm$^2$ the 6mm-diameter bore of the magnetic lenses becomes a strong limiting aperture for the resulting beam --- as is clearly seen in \ref{fig:pape} D and by the cut-off in the detected electron signal (YAG fluorescence).

@@

In principle, this unexpected increase in transverse momentum could be caused by any of three effects.
The first, and most easily ruled out, is that the electric field of the laser itself could affect the electrons, causing the divergence.
The possibility is easily dismissed as the effect is not observed when off the grating.
The second possible cause is the divergence of the plasmon's own electric field ($\partial E_{plasmon} / \partial x$).
This field is parallel to the surface plasmon vector $k_{SP}$ (assumed here to be in the $x$ direction).
The third possible cause is local field enhancement of the accelerating electric field due to the peaks in the sinusoidal shape of the surface, commonly known as the ``lightning rod'' effect.
For the employed one dimensional grating, this effect is necessarily parallel to the grating vector $k_G$.

@@

The experimental results are shown to the right in \ref{fig:rotated}.
As the divergence is clearly shown to be rotated by $45^{\circ}$ with respect to the colinear case, this experiment demonstrates that the cause of the divergence increase is due to the local field enhancements near the peaks of the grating.
This result is commensurate with the results shown by Zawadzka et al \cite{zawadzka_evanescent_2001}, who showed a reduced divergence in the Kretschmann geometry.
As that geometry does not employ a patterned surface, it would not be subject to such local field enhancements of the accelerating electric field.

@@

Other nano-engineered surfaces are still being considered for future work.
Various other more exotic nano-structures are under examination for PAPE \cite{li_surface_2013}.
Bi-metallic layers are known to enhance PAPE, which may allow for reduced incident laser power for driving the plasmon \cite{kupersztych_anomalous_2005}.
Photoemission from nano-arrays of quantum-wells also offers the tantalizing hope of reduced rms transverse momentum of the generated electrons.

>>>>

Above an incident laser pulse intensity of $\sim$32MW/cm$^2$, there is a rapid and dramatic increase in the emission efficiency for the nano-patterned photocathode.
This can be explained by a suppression of the barrier to photoemission below the two-photon excitation energy, thus allowing direct two-photon photoemission.
The well-known Schottky Effect states that for an applied electric field $E$ the work function will be suppressed by an amount
\begin{equation}
  \Delta \Phi = - e\sqrt{\frac{e E}{4 \pi \varepsilon_{\smallzero}}} \,\text{.}
\end{equation}
This suppression is likely caused by the laser-excited surface plasmon, since on the neighboring flat gold surface, where the suppression due to the laser (20MV/m) and the 20kV acceleration potential (0.8MV/m) is $\Delta \Phi \approx $ 0.18eV, no similar increase in efficiency is observed.
If this is indeed the case, and the surface plasmon field futher suppresses the barrier to a total $ \Delta \Phi \approx $ 0.35eV to allow direct two-photon photoemission, the field must be of the order of 60MV/m or three times that of the laser field.
In \ref{fig:pape} B, at a laser intensity of $\sim$30MW/cm$^2$, there is already a distinct broadening of the momentum distribution in the direction parallel to the wavevector $k_G$ of the sinusoidal modulation on the photocathode (i.e., perpendicular to the FIB-milled grooves). 
This broadening becomes the dominant feature of the electron emission above $\sim$32MW/cm$^2$; indeed, it becomes so severe that at laser intensities above $\sim$37MW/cm$^2$ the 6mm-diameter bore of the magnetic lenses becomes a strong limiting aperture for the resulting beam --- as is clearly seen in \ref{fig:pape} D and by the cut-off in the detected electron signal (YAG fluorescence).

@@

In principle, this unexpected increase in transverse momentum (\ref{fig:pape} B and C) could be caused by any of three effects.
The first, and most easily ruled out, is that the electric field of the laser itself could affect the electrons, causing the divergence.
The possibility is easily dismissed as the effect is not observed when off the grating.
The second possible cause is the divergence of the plasmon's own electric field ($\partial E_{plasmon} / \partial x$).
This field is parallel to the surface plasmon vector $k_{SP}$ (assumed here to be in the $x$ direction).
The third possible cause is local field enhancement of the accelerating electric field due to the peaks in the sinusoidal shape of the surface, commonly known as the ``lightning rod'' effect.
For the employed one dimensional grating, this effect is necessarily parallel to the grating vector $k_G$.

@@

The experimental results are shown to the right in \ref{fig:rotated}.
As the divergence is clearly shown to be rotated by $45^{\circ}$ with respect to the colinear case, this experiment demonstrates that the cause of the divergence increase is due to the local field enhancements near the peaks of the grating.
For a mathematical treatment of the transverse electric field of such patterned surfaces, the reader is referred to Ref. \cite{watts_sharp_1997}. 
Our results are commensurate with the results shown by Zawadzka et al \cite{zawadzka_evanescent_2001}, who showed a reduced divergence in the Kretschmann geometry.
As that geometry does not employ a patterned surface, it would not be subject to such local field enhancements of the accelerating electric field.

@@

Other nano-engineered surfaces are still being considered for future work.
Various other more exotic nano-structures are under examination for PAPE \cite{li_surface_2013,polyakov_plasmon_2013}.
Bi-metallic layers are known to enhance PAPE, which may allow for reduced incident laser power for driving the plasmon \cite{kupersztych_anomalous_2005}.
Photoemission from nano-arrays of quantum-wells also offers the tantalizing hope of reduced rms transverse momentum of the generated electrons.

### 2013-04-09 inc/photocathode/single_photon/single_photon.tex
more tweaks to the photocathode sections
Merge remaining comments from andreas on PAPE

\section{Single-photon Photoemission from Flat Metal Photocathodes}

Single-photon photoemission is the most common photoemission process and the easiest to explain.
As described by Einstein in his famous paper on ``the photoelectric effect'' \cite{einstein_uber_1905}, as it is colloquially known, if an electron absorbs a photon whose energy is sufficient to promote the electron above the local vacuum level, it become free to leave the host material.
The material dependent minimum energy needed to eject an electron is termed its ``work function'' $\Phi$.
Any additional energy above the work function remains with the ejected electron, usually in the form of kinetic energy, and is called the ``excess photoemission energy.''

@@

An example of single-photon photoemission is presented in \ref{fig:single_photon_tungsten}, wherein a sample of polished tungsten (W), having a work function $\Phi \approx $ 4.3-4.5 eV \cite{yen_thermally_1980}, is used as a photocathode.
When paired with our $\hbar \omega = $ 4.75 eV photon energy laser (see Section \ref{sec:laser}) the excess photoemission energy $\Delta E = \hbar \omega - \Phi \approx $ 0.35 eV.
The linear dependence of the electron yield (circles) on the laser pulse energy is indicative of a single photon process, as expected since each absorbed photon may eject one electron.
In this experiment, the yield is determined by integrating the total counts recorded on the CCD detector and therefore is only a relative measurement.
The fact that the Fourier plane spot-size (squares), directly related the initial transverse momentum variance $\Delta p_{\smallT}$, is unchanged over the full range of laser power is also expected, since by $\Delta p_{\smallT} = \sqrt{ m \Delta E / 3 } $ \cite{dowell_quantum_2009}, $\Delta p_{\smallT}$ only depends on $\Delta E$.

In order to improve performance, one must be able to minimize $\Delta p_{\smallT}$ without negatively affecting the photoemission efficiency, and thus electron yield.
As it stands currently, and as we have already seen in Section \ref{sec:must_reduce_transverse_momentum}, flat metal photocathodes employing single-photon photoemission are not likely to be useful for UEM since the transverse momentum variance cannot be reduced without also reducing the emission efficiency which is proportional to $\Delta E^2$ \cite{shalaev_electron_1994}.
However, as I will show in Section \ref{sec:photocathode_future}, this does not quite tell the complete story.

>>>>

\section{Single-photon Photoemission from Flat Metal Photocathodes} \label{sec:single_photon}

Single-photon photoemission is the most common photoemission process and the easiest to explain.
As described by Einstein in his famous paper on ``the photoelectric effect'' \cite{einstein_uber_1905}, as it is colloquially known, if an electron absorbs a photon whose energy is sufficient to promote the electron above the local vacuum level, it become free to leave the host material.
The material dependent minimum energy needed to eject an electron is termed its ``work function'' $\Phi$.
Any additional energy above the work function remains with the ejected electron, usually in the form of kinetic energy, and is called the ``excess photoemission energy'' $\Delta E$, for single photon photoemission, using a laser of frequency $\omega$ (photon energy $\hbar \omega$), this is simply,
\begin{equation}
  \Delta E = \hbar \omega - \Phi \,\text{.}
\end{equation}
The efficiency of photoemission $\eta_{{\scriptscriptstyle PE}}$, when $\Delta E$ is small compared to $\Phi$ may be written as,
\begin{equation}
  \eta_{{\scriptscriptstyle PE}} = A ( 1 - R ) \Delta E^2 \,\text{,}
\end{equation}
where $A$ is a material paramenter, and R is the photocathode surface reflectivity.

@@

An example of single-photon photoemission is presented in \ref{fig:single_photon_tungsten}, wherein a sample of polished tungsten (W), having a work function $\Phi \approx $ 4.3-4.5 eV \cite{yen_thermally_1980}, is used as a photocathode.
When paired with our $\hbar \omega = $ 4.75 eV photon energy laser (see Section \ref{sec:laser}) the excess photoemission energy $\Delta E \approx $ 0.35 eV.
The linear dependence of the electron yield (circles) on the laser pulse energy is indicative of a single photon process, as expected since each absorbed photon may eject one electron.
In this experiment, the yield is determined by integrating the total counts recorded on the CCD detector and therefore is only a relative measurement.
The fact that the Fourier plane spot-size (squares), directly related the initial transverse momentum variance $\Delta p_{\smallT}$, is unchanged over the full range of laser power is also expected, since by $\Delta p_{\smallT} = \sqrt{ m \Delta E / 3 } $ \cite{dowell_quantum_2009}, $\Delta p_{\smallT}$ only depends on $\Delta E$.

In order to improve performance, one must be able to minimize $\Delta p_{\smallT}$ without negatively affecting the photoemission efficiency, and thus electron yield.
For single-photon photoemission, although the polarization-dependent reflectivity can be tweaked to improve performance slightly \cite{berger_dc_2009}, due to the quadratic dependence of the efficiency on excess photoemission energy \cite{shalaev_electron_1994}, any reduction in $\Delta p_{\smallT}$ gained by reducing $\Delta E$ will be mitigated by a large drop in electron yield.
For this reason, as it stands currently and as we have already seen in Section \ref{sec:must_reduce_transverse_momentum}, flat metal photocathodes employing single-photon photoemission are not likely to be useful for UEM.
However, as I will show in Section \ref{sec:photocathode_future}, this does not quite tell the complete story.
